---
title: "Further Analysis of all BBT Messages, Focus on keywords: Behavior(s), Milestones, Toddler(s)"
author: "Kat"
date: "1/11/2018"
output: 
  html_document:
    pdf_document:
      keep_md: true
---

* Note: Spelling, Stems, and Plurals will look wrong because data was cleaned to remove punctuation and numbers.

* 0.1 Total & Omitted Messages

* 1.1 Total Number of "Behavior(s)" Messages
* 1.2 Total Number of "Behavior(s)" Messages with CTR >= 15%
* 1.3 Total Number of "Behavior(s)" Messages with CTR < 15%

* 2.1 Total Number of "Milestones" Messages
* 2.2 Total Number of "Milestones" Messages with CTR >= 15%
* 2.3 Total Number of "Milestones" Messages with CTR < 15%

* 3.1 Total Number of "Toddler(s)" Messages
* 3.2 Total Number of "Toddler(s)" Messages with CTR >= 15%
* 3.3 Total Number of "Toddler(s)" Messages with CTR < 15%

* 4.1 Total Number of "Babies" Messages
* 4.2 Total Number of "Babies" Messages with CTR >= 15%
* 4.3 Total Number of "Babies" Messages with CTR < 15%

* 5.1 Total Number of "Fun" Messages
* 5.2 Total Number of "Fun" Messages with CTR >= 15%
* 5.3 Total Number of "Fun" Messages with CTR < 15%

* 6.1 Total Number of "Development" Messages
* 6.2 Total Number of "Development" Messages with CTR >= 15%
* 6.3 Total Number of "Development" Messages with CTR < 15%

* 7.1 Total Words in Messages with CTR >=15%, stop words removed
* 7.2 Total Words in Messages with CTR <7%, stop words removed
* 7.3 Words Unique to CTR >=15%

* 8.1 Milestone Messages Over Time

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r, message=FALSE, echo=FALSE}
# First, I loaded in the Data and took out empty rows at the bottom of the spreadsheet. (see .rmd file for code)
rm(list = ls())
options(java.parameters = "-Xmx4g")
library(rJava)
library(xlsx)
library(dplyr)
library(tidytext)
library(ggplot2)
library(stringi)
library(SnowballC)
library(tidyverse)
library(gdata)

files <- list.files("BBT Message Library 1.9.2018", pattern = "*.xlsx")
for (i in seq_along(files)) {
    wb <- loadWorkbook(...)
    ...
    rm(wb)
    xlcFreeMemory()  # <= free Java Virtual Machine memory !
}


bbt <- read.xlsx2("BBT Message Library 1.9.2018.xlsx", sheetIndex=1, header=TRUE)

```

```{r, echo=FALSE}
# Read in Tagged Messages
files <- list.files("taggedmessages8", pattern = "*.xlsx")
for (i in seq_along(files)) {
    wb <- loadWorkbook(...)
    ...
    rm(wb)
    xlcFreeMemory()  # <= free Java Virtual Machine memory !
}
bbt_tagged <- read.xlsx("taggedmessages.xlsx", sheetIndex=1, header=TRUE)
```

#### 0.1 Total & Omitted Messages
- There were originally 630 messages
- 10 onboarding messages were removed (Month <0), leaving 620 for analysis
```{r}
# Convert to UTF8, subset required columns
bbt$English.Message <- iconv(bbt$English.Message, "UTF-8", "UTF-8")
bbt_tagged$English.Message <- iconv(bbt_tagged$English.Message, "UTF-8", "UTF-8")
#nrow(bbt) 630 original rows
# for the rest of bbt_tagged see section 5

#subset required columns
bbt2 <- bbt[,c(1, 2, 3, 5, 6, 7)] # Keep Month, Day, English.Message, English.Link.Views, English.Link.Ctr, English.Title

#subset messages with month <0 for BBT (on boarding messages)
bbt2$Month <- as.numeric(levels(bbt2$Month))[bbt2$Month]
bbt2$English.Link.CTR <- as.numeric(levels(bbt2$English.Link.CTR))[bbt2$English.Link.CTR]
onboarding <- subset(bbt2, Month <0)
onboarding$English.Message

#dataset exluding onboarding messages
bbt2 <- subset(bbt2, Month >=0)  #10 rows removed as onboarding messages
rm(bbt)
```

```{r}
#colnames(bbt2)
```

```{r}
bbt3<- as.data.frame(bbt2)
bbt3$English.Message <- gsub("BBT: ", "", bbt3$English.Message)
# milestones messages without numbers removed
p <- grep("milestones", bbt3$English.Message, ignore.case=TRUE)
milestone=NULL
for (i in p) {
  milestone=rbind(milestone,(bbt3[i,]))
} 
#e[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
#milestone$English.Message

```


```{r, echo=FALSE}
#convert to data frame, fit typos
bbt3<- as.data.frame(bbt2)
bbt3$English.Message <- gsub("BBT: ", "", bbt3$English.Message)#subset out ignoring messages above & remove BBT: in text.message
bbt3$English.Message <- gsub("'s", "", bbt3$English.Message) # remove 's
bbt3$English.Message <- gsub("[^[:alpha:]]", " ", bbt3$English.Message)

```

### 1.1 Total Messages with "Behavior(s)", "Milestones" and "Toddler(s)" 
- 19 total messages with behavior/behaviors
```{r}

p <- grep("behaviors|behavior", bbt3$English.Message, ignore.case=TRUE)
f=NULL
for (i in p) {
  f=rbind(f,(bbt3[i,]))
}
#f[, c(1,3,5)]   #to double check accuracy of subset & variables with CTR
f$English.Message

```
### 1.2 Messages with "Behavior(s)" with Click-Through-Rate(CTR) greater than 15%
- 182 Messages have CTR >=15%
- 12 of the 19 "behavior(s)" messages have CTR greater than 15%
```{r}
bbt3High <- subset(bbt3, English.Link.CTR >= 0.15)
p <- grep("behavior|behaviors|Behavior|Behaviors", bbt3High$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3High[i,]))
} 
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
g$English.Message

```

### 1.3 Messages with "Behavior(s)" with Click-Through-Rate(CTR) less than 15%
- 7 of the 19 "behavior(s)" messages have CTR less than 15%, but the CTR rates were only a little lower with a range from 12-14%
```{r}
bbt3Less <- subset(bbt3, English.Link.CTR < 0.15)
p <- grep("behavior|behaviors|Behavior|Behaviors", bbt3Less$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3Less[i,]))
} 
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
#mean(g$English.Link.CTR)
g$English.Message
```


#### 2.1 Total messages with word Milestone - 8 total messages with word milestones
- 8 total messages with milestones
```{r}

p <- grep("milestones", bbt3$English.Message, ignore.case=TRUE)
e=NULL
for (i in p) {
  e=rbind(e,(bbt3[i,]))
} 
#e[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
#e$English.Message
milestone$English.Message #to include all #s
```

#### 2.2 Milestones Messages with Click-Through-Rate(CTR) greater than 15%
- 182 Messages have CTR >=15%
- All "Milestones" messages have a CTR greater than 15%
```{r}
p <- grep("milestones", bbt3High$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3High[i,]))
}
#g[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
#g$English.Message
milestone$English.Message # to include #s
```
#### 2.3 Milestones Messages with Click-Through-Rate(CTR) less than 15%
- All "Milestones" messages have a CTR less than 15% (so none in this section)
```{r}
p <- grep("milestones", bbt3Less$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3Less[i,]))
} 
#g[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
g$English.Message
```

#### 3.1 Total messages with word Toddler/s - 36 total messages with word toddler/toddlers
- 36 total messages with toddler(s)
```{r}
p <- grep("toddler|toddlers|Toddler|Toddlers", bbt3$English.Message, ignore.case=TRUE)
d=NULL
for (i in p) {
  d=rbind(d,(bbt3[i,]))
} 
#d[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
d$English.Message
```

#### 3.2 Toddler(s) Messages with Click-Through-Rate(CTR) greater than 15%
- 182 Messages have CTR >=15%
- 24 of the 36 "toddler(s)" messages have CTR greater than 15%
```{r}
p <- grep("toddler|toddlers|Toddlers|Toddler", bbt3High$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3High[i,]))
}
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
g$English.Message
```
#### 3.3 Toddler(s) Messages with Click-Through-Rate(CTR) less than 15%
- 12 of the 36 "toddler(s)" messages have CTR less than 15%. The range of the CTR for those 12 messages is 7-14%, with an average of 11.6 and median of 12.
```{r}
p <- grep("toddler|toddlers|Toddlers|Toddler", bbt3Less$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3Less[i,]))
} 
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
#mean(g$English.Link.CTR)
#median(g$English.Link.CTR)
g$English.Message
```

#### 4.1 Total Message with word "babies"
- 14 messages with babies
- 92 messages with baby
```{r, echo=FALSE}
p <- grep("babies|baby", bbt3$English.Message, ignore.case=TRUE)
a <- grep("baby", bbt3$English.Message, ignore.case=TRUE)
b <- grep("baby's", bbt3$English.Message, ignore.case=TRUE)
d=NULL
for (i in p) {
  d=rbind(d,(bbt3[i,])) #babies
} 
e=NULL
for (i in a) {
  e=rbind(e,(bbt3[i,])) #baby
}

f=NULL
for (i in b) {
  f=rbind(f,(bbt3[i,])) #baby's
}

#d[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
#length(d$English.Message) #14 messages with "babies"
#length(e$English.Message) # 92 messages with "baby"
#length(f$English.Message) # 0 messages with baby's b/c I took out 's
#d[!(d %in% e)]
d$English.Message        
```
#### 4.2 Babies Messages with Click-Through-Rate(CTR) greater than 15%
- 182 Messages have CTR >=15%
- 12 of the 14 "babies" messages have CTR greater than 15%
```{r}
p <- grep("babies", bbt3High$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3High[i,]))
}
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
g$English.Message

```
#### 4.3 Babies Messages with Click-Through-Rate(CTR) less than 15%
- There are 2 messages with the word "babies" with CTR less than 15% ; one message has no Link, the other is pre-natal message
```{r}
p <- grep("babies", bbt3Less$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3Less[i,]))
} 
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
mean(g$English.Link.CTR)
median(g$English.Link.CTR)
g
```

#### 5.1 Total Messages with word "fun"
- 50 total messages with "fun"
```{r, echo=FALSE}
p <- grep("fun", bbt3$English.Message, ignore.case=TRUE)
d=NULL
for (i in p) {
  d=rbind(d,(bbt3[i,])) #fun
} 
d$English.Message
```
#### 5.2 Fun Messages with Click-Through-Rate(CTR) greater than 15%
- 182 Messages have CTR >=15%
- 15 of the 50 "fun" messages have CTR greater than 15%
- The mean age a >15% "fun" message is sent is 17 months. The median age is 15 months.
```{r}
p <- grep("fun", bbt3High$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3High[i,]))
}
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
g$English.Message
#mean(g$Month)
#median(g$Month)

```
#### 5.3 Fun Messages with Click-Through-Rate(CTR) less than 15%
- 35 of the 50 messages with the word "fun" have CTR <15% - this is the majority of messages!
- The mean and median of a "fun" message is 10%.
- The mean age that a "fun" message is sent is 33 months. The median age is 36. (Compared to 17/15 for "fun" messages with CTR >15%)
- Possibly, users are not clicking on the message because it says "fun" but rather a few fun messages are sent earlier in the child's life.
```{r}
p <- grep("fun", bbt3Less$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3Less[i,]))
} 
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
#mean(g$Month)
#median(g$Month)
g$English.Message
```

#### 6.1 Total Messages with word "development"
- 27 total messages with "development"
```{r, echo=FALSE}
p <- grep("development", bbt3$English.Message, ignore.case=TRUE)
d=NULL
for (i in p) {
  d=rbind(d,(bbt3[i,])) #development
} 
d$English.Message
```

#### 6.2 Development Messages with Click-Through-Rate(CTR) greater than 15%
- 6 of the 14 (>15%) Development messages also contain "milestones" 
```{r}
p <- grep("development", bbt3High$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3High[i,]))
}
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
g$English.Message

```

#### 6.3 Development Messages with Click-Through-Rate(CTR) less than 15%
- 13 of the 27 development messages have CTR <15%
- But, development messages still have a mean CTR of 12% and median of 14%. This is probably because many of the milestones messages contain the word development.
- Mean age of message sent is 33 months. Median age is 40 months.
```{r}
p <- grep("development", bbt3Less$English.Message, ignore.case=TRUE)
g=NULL
for (i in p) {
  g=rbind(g,(bbt3Less[i,]))
} 
#g[, c(1,3,5)] #to double check accuracy of subset & variables with CTR
#mean(g$English.Link.CTR)
#median(g$English.Link.CTR)
g$English.Message
```


#### 7.1 There are 42 total words (with stop words removed) from messages with CTR >= 15%
```{r, message=FALSE}
#bbt3High >=15%
bbt_dfH <- data_frame(line=nrow(bbt3High), text=bbt3High$English.Message, CTR=bbt3High$English.Link.CTR)

## use tidytext to unnest all Text messages per line,  unest words, and unjoin stop words (INCLUDES ALL)
bbt_dfH <- bbt_dfH %>% unnest_tokens(word, text) %>% anti_join(stop_words) %>%
  count(word, sort=TRUE) %>% 
  filter(n>5) %>%  
  mutate(word=reorder(word, n))
#bbt_dfH
```

#### 7.2 There are 29 total words (with stop words removed) from messages with CTR <= 7% (Lowest CTR rates)
```{r, message=FALSE}
bbt3Low <- subset(bbt3, English.Link.CTR <= 0.07)
bbt_dfL <- data_frame(line=nrow(bbt3Low), text=bbt3Low$English.Message, CTR=bbt3Low$English.Link.CTR)

## use tidytext to unnest all Text messages per line,  unest words, and unjoin stop words (INCLUDES ALL)
bbt_dfL <- bbt_dfL %>% unnest_tokens(word, text) %>% anti_join(stop_words) %>%
  count(word, sort=TRUE) %>% 
  filter(n>5) %>%  
  mutate(word=reorder(word, n))
#bbt_dfL

```
#### 7.3 Unique Words to CTR >= 15%
- There are 10 words that are common to both CTR >=15% and CTR <=7% (child    learn    baby     tips     time     talk     teeth    children reading  healthy)
- There are 32 unique words to CTR >=15% (Ignore "42 Levels" below)
```{r}
#####What are Common words from CTR <= 7% and CTR >=15 %?
wordStop15H<- bbt_dfH$word # words from high CTR
wordStop15L <- bbt_dfL$word # words from low CTR

JointWORD_stop <- wordStop15H[wordStop15H %in% wordStop15L] # what high words are in low words?

uniqueWORD_stop <- wordStop15H[!(wordStop15H %in% JointWORD_stop)]
uniqueWORD_stop
```


```{r}
#### 5.1 Milestone Messages Over Time  - updated with both types of messages below
# p <- grep("milestones", bbt3High$English.Message, ignore.case=TRUE)
# g=NULL
# for (i in p) {
#   g=rbind(g,(bbt3High[i,]))
# }
# #g[, c(1,3,5)]  #to double check accuracy of subset & variables with CTR
# #g$English.Message
# milestone$English.Message #to include #s in milestone messages for references
# 
# ggplot(data=g, aes(x=Month, y=English.Link.CTR))+
#   geom_point() +
#   ggtitle("Are Users Clicking on Milestone Messages Less As Their Child Ages?")+
#   xlab("Child's Age in Months")+
#   ylab("Click-Through Rate Per Message")

```

#### 8.1 Milestone Message Over Time
- 19 total messages identified by Lynn as a "Milestones" Message
- From analysis above, 8 of these message have the word "milestones" (aqua) included in the text message, leaving 11 messages identified as milestone message without have "Milestones" in the actual text message (coral)
- Aqua "milestone2" messages are those that HAVE the word "milestones" in the actual text message
- Coral "milestone" messages are those that DO NOT have the word "milestones" in the actual text message
- Milestone2(aqua) messages have been around longer and have Click Views ranging from 398 to 724, while Milestone (coral) messages have a range of 51 to 610. 
- Of Milestone(coral) labeled messages, the highest Views and CTRs are when the child is aged 23, 26, 32 with Views of 610, 455, and 321, and CTR of 24%, 18%, and 15%.
- The stats on the remaining 8 Milestone(coral) messages are: Views ranging from 29-80, and CTR ranging from 7-14%. These may grow as subscribers's children reach these older ages.

```{r}
bbt_tagged2 <- bbt_tagged[,c(1, 2, 3, 4, 6, 7, 8)]  #removing necessary columns
#bbt_tagged2

# finding messages with the word milestone in them
   p <- grep("milestones", bbt_tagged2$English.Message, ignore.case=TRUE)
milestone=NULL
for (i in p) {
  milestone=rbind(milestone,(bbt_tagged2[i,]))
}
# assign specific type to milestone word messages
milestone$Type <- "milestone2"
#milestone #milestone worded messages only

#hack way of combining above with all message - could not figure out how to do if/then statement 
# filter out milestone & milestone2, plot results
whole <- rbind(bbt_tagged2, milestone)  # combine all with milestone worded messages only
milestones <- whole[9:53,]  #remove unlabeled messages
milestones <- as.data.frame(milestones) # filter out only milestone messages
milestones %>%
  filter(Type == "milestone2" | Type =="milestone") %>%
  ggplot(aes(x=Month, y=English.Link.CTR, color=Type))+
  geom_point() + 
  ggtitle("CTR for Milestone Identified & Labeled Messages As Child Ages")+
  xlab("Child's Age in Months")+
  ylab("CTR")+
  scale_fill_discrete(labels=c("No Milestones in Message", "Milestones in message"))

milestone$English.Message  # for reference with graph
```


